Lipid Interactomics Repository

### 2025-02-03 BioinformaticsApplicationNote/GG_2025_applicationNote.qmd
Minor updates

---
title: "Lipid Interactomics Repository"
subtitle: "An interactive and open access platform for exploring multifunctional lipid interactomes"
authors:
  - name: "Fikadu Tafesse"
    affiliation: "Oregon Health & Science University"
    orcid: 0000-0002-8575-4164
  - name: "Carsten Schultz"
    affiliation: "Oregon Health & Science University"
    orcid: 0000-0002-5824-2171
  - name: "Gaelen Guzman"
    affiliation: "Oregon Health & Science University"
    orcid: 0000-0002-7696-6034
    corresponding: true
    email: gaelenguzman@lipidinteractome.org
format:
  html:
    embed-resources: true
  docx:
    fig-height: 5
    reference-doc: custom-reference-doc.docx
page-navigation: false

@@

- Description
  * A site built using [Quarto](https://quarto.org)
  * FAIR data principles
    - *Findable:* We sought to create a platform to centralize the studies of this type. The site is organized such that one can interact with the results of individual published studies or explore the combined results of all studies which investigated a given lipid. Because this site is a living document, we invite researchers to contribute their results by contacting us with the details delineated at our [Data Submission](/ContactUs/DataSubmission.qmd) page. 
    - *Accessible:* To enhance the accessibility of these data, we have made the results of all studies and lipid probes downloadable in .csv format. We have provided a [detailed overview of the data production and analysis](/Background/ProteomicsUsingMultifunctionalProbes.qmd), a description of the content in all columns, and a [nuanced discussion about the limitations of these probes](/Background/MultifunctionalLipidProbesOverview.qmd). To maximize the reliability of each reported interactome, we have focused on amassing only peer-reviewed proteomics results.
    - *Interoperable:* We have sought to coordinate the formatting of the data collected from a variety of research teams. While many studies utilized identical data analysis pipelines, some of our included publications used distinct mass spectrometry protocols and we have attempted to provide a parallel data structure and visualization -- while also [explaining the differences in methodology](/Background/ProteomicsUsingMultifunctionalProbes.qmd).
    - *Reusable:* We have provided extensive documentation on the production, analyses, and caveats of these data; in these documentations, we have sought to limit technical language and to give these results an accessible interpretation. Additionally, we have referenced numerous publications which go into much greater depth on the synthesis of these lipid probes and the methodology used in each proteomics analysis. 
  * An interactive discovery platform intended to guide lipid research
  * Contains data from 5 studies thus far, though framework for expanding to many more.
  * Enables the direct comparison of different studies, and is an invaluable addition to the understanding of how proteins interact with lipids

@@

    - Where similar study design allows, data is incorporated into a Shiny app for interactive selection of probes for exploration at [Probe vs Probe Comparisons](/LipidProbe/EnrichedHitsComparison.qmd). These plots incorporate linear regression analysis to describe the similarity of the datasets, and highlight proteins which were significantly enriched in one or the other dataset -- or both datasets.
  * The site is a living document and is expected to grow as future studies utilize these functionalized lipid analogs.

- Acknowledgements

>>>>

---
title: "Lipid Interactomics Repository"
subtitle: "An interactive and open access platform for exploring multifunctional lipid interactomes"
authors:
  - name: "Fikadu Tafesse"
    affiliations:
      - name:  "Oregon Health & Science University"
        address: 3222 SW Research Dr
        city: Portland
        state: OR
        postal-code: 97239
    orcid: 0000-0002-8575-4164
  - name: "Carsten Schultz"
    affiliations:
      - name:  "Oregon Health & Science University"
        address: 3222 SW Research Dr
        city: Portland
        state: OR
        postal-code: 97239
    orcid: 0000-0002-5824-2171
  - name: "Gaelen Guzman"
    affiliations:
      - name:  "Oregon Health & Science University"
        address: 3222 SW Research Dr
        city: Portland
        state: OR
        postal-code: 97239
    orcid: 0000-0002-7696-6034
    corresponding: true
    email: gaelenguzman@lipidinteractome.org
format:
  html:
    embed-resources: true
  docx:
    fig-height: 5
    reference-doc: /Users/gaelenguzman/LipidInteractomics_Website_local/LipidInteractomics_Website/Resources/custom-reference-doc.docx
page-navigation: false

@@

- Description
  * A site built using [Quarto](https://quarto.org). The R scripts which define the rendered figures is embedded in the source material for each page.
  * FAIR data principles
    - *Findable:* We sought to create a platform to centralize the studies of this type. The site is organized such that one can interact with the results of individual published studies or explore the combined results of all studies which investigated a given lipid. Because this site is a living document, we invite researchers to contribute their results by contacting us with the details delineated at our [Data Submission](/ContactUs/DataSubmission.qmd) page. 
    - *Accessible:* To enhance the accessibility of these data, we have made the results of all studies and lipid probes downloadable in .csv format. We have provided a [detailed overview of the data production and analysis](/Background/ProteomicsUsingMultifunctionalProbes.qmd), a description of the content in all columns, and a [nuanced discussion about the limitations of these probes](/Background/MultifunctionalLipidProbesOverview.qmd). To maximize the reliability of each reported interactome, we have focused on amassing only peer-reviewed proteomics results.
    - *Interoperable:* We have sought to coordinate the formatting of the data collected from a variety of research teams. While many studies utilized identical data analysis pipelines, some of our included publications used distinct mass spectrometry protocols and we have attempted to provide a parallel data structure and visualization -- while also [explaining the differences in methodology](/Background/ProteomicsUsingMultifunctionalProbes.qmd).
    - *Reusable:* We have provided extensive documentation on the production, analyses, and caveats of these data; in these documentations, we have sought to limit technical language and to give these results an accessible interpretation. Additionally, we have referenced numerous publications which go into much greater depth on the synthesis of these lipid probes and the methodology used in each proteomics analysis. 
  * An interactive discovery platform intended to guide lipid research
  * Contains data from 5 studies thus far, though framework for expanding to many more.
  * Enables the direct comparison of different studies, and is an invaluable addition to the understanding of how proteins interact with lipids

@@

    - Where similar study design allows, data is incorporated into a Shiny app for interactive selection of probes for exploration at [Probe vs Probe Comparisons](/LipidProbe/EnrichedHitsComparison.qmd). These plots incorporate linear regression analysis to describe the similarity of the datasets, and highlight proteins which were significantly enriched in one or the other dataset -- or in both datasets. We seek for this functionality to enable researchers to identify targets of interest when studying bioactive lipids. 
  * The site is a living document and is expected to grow as future studies utilize these functionalized lipid analogs. We have prepared a [Data Submission page](/ContactUs/DataSubmission.qmd) for any researchers with similar datasets who wish to contribute to this repository.

- Discussion
  *	What we hope this resource enables researchers to do
  *	Longevity – this site is designed to expand with the acquisition of this types of proteomics analyses

- Acknowledgements
  * Andre Nadler
  * Frank Stein
  * Per Haberkant
  *

- Conflicts of Interest
  * None

- References
